---
title: "Assignment_1_IRIS"
output:
  html_document:
    df_print: paged
  word_document:
    fig_width: 5
    fig_height: 5
    fig_caption: yes
---
#3.1 Reading and Manually Checking

#3.1a View the file in a text-editor to determine its format and read the file into R.

iris <- read.csv("dirty_iris.csv", stringsAsFactors = FALSE)
```{r}
iris <- read.csv("dirty_iris.csv", stringsAsFactors = FALSE)
```

#3.1b Calculate the number and percentage of observations that are complete.
# n is number of observations
```{r}
n <- complete.cases(iris)
percent <- (sum(n)/nrow(iris)) * 100
paste("The percentage of complete observation:",percent,"%")
```
#3.1c Does the data contain other special values? If it does, replace them with NA

```{r}
is.special <- function(x){
  if(is.numeric(x)) !is.finite(x)
  else is.na(x)
}
sapply(iris, is.special)

for(i in colnames(iris)){
  is.na(iris[[i]]) <- is.special(iris[[i]])
}
summary(iris)
head(iris)
```

#3.2 Checking with the rules

#3.2a Besides missing values, the data set contains errors. We have the following background knowledge:
# -Species should be one of the following values: setosa, versicolor or virginica
# -All measured numerical properties of an iris should be positive.
# -The petal length of an iris is at least 2 times its petal width.
# - The sepal length of an iris cannot exceed 30 cm.
# -The sepals of an iris are longer than its petals
#Define these rules in a separate text file and read them into R using editfile

```{r}
library(editrules)
E <- editfile("constraints_iris.txt")
print(E)
```
#3.2b Determine how often each rule is broken (violatedEdits). Summarize and plot the result.

```{r}
ve <- violatedEdits(E, iris)
summary(ve)
plot(ve)
```

#3.2c What percentage of the data has no errors

```{r}
percent_noerrors <- (sum(apply(ve, 1, sum)==0, na.rm = TRUE) / nrow(iris)) * 100
percent_noerrors_round <- round(percent_noerrors, digits = 1)
paste("The percentage of data that has no error:",percent_noerrors_round,"%")
```

#3.2d which observations have too long petals using the result of violatedEdits.
```{r}
observation <- which(iris$Petal.Length >= iris$Sepal.Length)
```

#3.2e Find outliers in sepal length using boxplot and boxplot.stats
```{r}
boxplot(iris$Sepal.Length)

outlier <- boxplot.stats(iris$Sepal.Length)$out
```
#setting outliers to NA
```{r}
new_iris <- iris
outlier_1 <- new_iris$Sepal.Length %in% outlier
new_iris$Sepal.Length[outlier_1] <- rep(NA, length(outlier))
boxplot(new_iris$Sepal.Length)
```
#3.3a Replace non positive values from Petal.Width with NA using correctWithRules from the library deducorrect.
```{r}
library(deducorrect)
cr <- correctionRules(expression(
  if (!is.na(Petal.Width) && Petal.Width <=0 ) Petal.Width = NA
))
new_iris <- correctWithRules(cr, new_iris)$corrected
```
#3.3b Replace all erronous values with NA using (the result of) localizeErrors
```{r}
head(localizeErrors(E, iris)$adapt)
new_iris[localizeErrors(E, new_iris)$adapt] <- NA
summary(violatedEdits(E, new_iris))
any(violatedEdits(E,new_iris), na.rm = TRUE)
```
#3.4 Imputing

#3.4a Use kNN imputation (VIM) to impute all missing values
```{r}
library(VIM)
iris_knn <- kNN(new_iris)
petal_width_knn <- iris_knn$Petal.Width
summary(petal_width_knn)
head(iris_knn[,names(new_iris)], 10)
```
#3.4b Use sequential hotdeck imputation to impute Petal.Width by sorting the dataset on Species. 
#Compare the imputed Petal.Width with the sequential hotdeck imputation method. 
#Note the ordering of the data!
```{r}
seqImp <- function(x, last=median(x, na.rm = TRUE)){
  n <- length(x)
  x <- c(x, last)
  j <- is.na(x)
  while (any(j)) {
    x[j] <- x[which(j)+1]
    j <- is.na(x)
  }
  x[1:n]
}
new_iris_So <- new_iris[order(new_iris$Species),]
head(new_iris_So)
new_iris_So$Petal.Width<- seqImp(new_iris_So$Petal.Width)
```

#3.4c Do the same but now by sorting the dataset on Species and Sepal.Length
```{r}
new_iris_Sl <- new_iris[order(new_iris$Species, new_iris$Sepal.Length),]
head(new_iris_Sl)
new_iris_Sl$Petal.Width<- seqImp(new_iris_Sl$Petal.Width)
```
#Comparing the dataframe
```{r}
comp_df <- data.frame(order1 = new_iris_So$Petal.Width, order2 = new_iris_Sl$Petal.Width)
head(comp_df)
```
#We suspect that order1 and order2 has the exact same values.
#Hence to confirm we perform setdiff and sort function
```{r}
setdiff(comp_df$order1, comp_df$order2)
sort(unique(comp_df$order1))
sort(unique(comp_df$order2))
```
